Designing a high-performance boundary element library with OpenCL and Numba

### 2021-05-11 opencl.tex
american spelling and add Numba reference

\begin{itemize}
	\item Test and trial indices of triangles over which to be integrated.
	\item Signs of the normal directions for the spaces.
	\item Test and trial grids as flat floating point arrays, defining each triangle through nine floating point numbers, specifying the $(x, y, z)$ coordinates of each of the three nodes of a triangle.
	\item Test and trial connectivity, which are lists of node indices that define the corresponding elements of the test and trial grid.
	\item Test and trial mappings of local triangle degrees of freedom to global degrees of freedom.
	\item Test and trial basis function multipliers for each triangle, which are triangle dependent prefactors needed for certain function spaces (e.g. in electromagnetics).
	\item Quadrature points and quadrature weights.
	\item A buffer that contains the global assembled matrix.
	\item Additional kernel parameters, such as the wavenumber for Helmholtz problems.
	\item The number of test and trial degrees of freedom.
	\item A single byte that is set to one if the test and trial grids are disjoint.
\end{itemize}

@@

\begin{figure}
	\center
	\input{img/colouring}
	\caption{The triangles in this mesh have been colored (using a greedy algorithm) so that no two neighbouring triangles are the same color. Sets of triangles of the same color can therefore be processed together as they are guaranteed to not be neighbors, so do not share any degrees of freedom. [The dots on each cell are included as a visual aid for anyone who prints this article in black and white.]}
	\label{fig:colouring}
\end{figure}

\subsection{Avoiding data races in the global assembly}
Data races in global assembly routines are a problem whenever different triangles need to sum into the same global degree of freedom. To solve this we use standard colouring techniques to split up the computations into chunks that access different data regions. To this end, we define two triangles as neighbours if they share at least one global degree of freedom. Based on this relationship we run a simple greedy colouring algorithm in the initialization phase of a function space. An example colouring is shown in \cref{fig:colouring}.

The compute kernels can then be run colour-by-colour. OpenCL parallelizes over test elements and trial elements, and so we have to iterate over the product space of possible colour combinations. In Numba, we only parallelise over test elements so it is sufficient to iterate over all possible colours in the test space.

>>>>

\begin{itemize}
	\item Test and trial indices of triangles over which to be integrated.
	\item Signs of the normal directions for the spaces.
	\item Test and trial grids as flat floating point arrays, defining each triangle through nine floating point numbers, specifying the $(x, y, z)$ coordinates of each of the three nodes of a triangle.
	\item Test and trial connectivity, which are lists of node indices that define the corresponding elements of the test and trial grid.
	\item Test and trial mappings of local triangle degrees of freedom to global degrees of freedom.
	\item Test and trial basis function multipliers for each triangle, which are triangle dependent prefactors needed for certain function spaces (e.g., in electromagnetics).
	\item Quadrature points and quadrature weights.
	\item A buffer that contains the global assembled matrix.
	\item Additional kernel parameters, such as the wavenumber for Helmholtz problems.
	\item The number of test and trial degrees of freedom.
	\item A single byte that is set to one if the test and trial grids are disjoint.
\end{itemize}

@@

\begin{figure}
	\center
	\input{img/colouring}
	\caption{The triangles in this mesh have been colored (using a greedy algorithm) so that no two neighboring triangles are the same color. Sets of triangles of the same color can therefore be processed together as they are guaranteed to not be neighbors, so do not share any degrees of freedom. [The dots on each cell are included as a visual aid for anyone who prints this article in black and white.]}
	\label{fig:colouring}
\end{figure}

\subsection{Avoiding data races in the global assembly}
Data races in global assembly routines are a problem whenever different triangles need to sum into the same global degree of freedom. To solve this we use standard coloring techniques to split up the computations into chunks that access different data regions. To this end, we define two triangles as neighbors if they share at least one global degree of freedom. Based on this relationship we run a simple greedy coloring algorithm in the initialization phase of a function space. An example coloring is shown in \cref{fig:colouring}.

The compute kernels can then be run color-by-color. OpenCL parallelizes over test elements and trial elements, and so we have to iterate over the product space of possible colour combinations. In Numba, we only parallelise over test elements so it is sufficient to iterate over all possible colours in the test space.
